---
title: "Introduction to geopr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to geopr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r }
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## What is geopr?

The **geopr** package provides programmatic access to official geospatial
datasets for Puerto Rico published through web services, with a particular
focus on open standards such as WFS (Web Feature Service).

Its primary goal is to reduce friction between R users and institutional
geospatial services by enabling reproducible, transparent exploration and
download of spatial data.

## What is a WFS?

A Web Feature Service (WFS) is an Open Geospatial Consortium (OGC) standard
that allows access to vector-based geographic features over the web.

Unlike map services (WMS), a WFS returns complete geographic data
(geometry and attributes), making it suitable for direct analysis in R
using packages such as `sf`.

## Basic workflow in geopr

A typical **geopr** workflow consists of two main steps:

1. Explore available layers in the service.
2. Download a specific layer for analysis.

```{r }
library(geopr)
```

### 1. Explore available layers
# \dontrun{
layers <- geopr_list_layers()
head(layers)
# }

The result is a table listing technical layer names and basic metadata,
allowing users to identify which datasets are relevant for their analysis.

### 2. Download a layer

```{r}
# \dontrun{
municipalities <- geopr_get_layer("pr_geodata:g03_legales_municipios_2023")
municipalities
# }
```

The returned object is an `sf` object, ready for spatial analysis,
visualization, or integration with statistical data.

## Important considerations

- WFS services depend on external infrastructure and may experience outages
  or changes.
- Technical layer names may change over time.
- geopr prioritizes simplicity and reproducibility over exposing all
  available metadata.



